\chapter{Discussion}\label{sec:Discussion}

\section{Inverted Pendulum Possible Improvements}

The inverted pendulum fulfills all requirements and the possible simple improvements such as the suppression of the shaking are already discussed in \autoref{sec:InvPendImp}. What could have been done however, is to try other control design schemes such as state space equations or bode plots and test them to see which one performs the best. Unfortunately too much time was spent on fixing the setup to be able to do so.

\section{Rocket}

The rocket as it is has not been launched. Nevertheless the rocket could be improved in its physical design and on the controller chosen. As explained in \autoref{sec:RocketControllerDesign} a controller similar to the Inverted Pendulum controller would be found by having an inner loop controlling the servomotors. The inertia model used was very simplified, a more complex one, e.g adding the rocket's lighter parts in the equations, could improve the performance of the controller. Similarly to the Inverted Pendulum improvements, other control design schemes, such as space equations or bode plot, could be implemented. It is to be noted that a lot time was spent on the rocket design and construction. On future implementations a ready-made rocket could be used to spare time.

\section{Differences between the Rocket and the Inverted Pendulum}

As explained in \autoref{sec:ModelComp} the models are slightly different. First of all, opposite angles were chosen in the beginning of the modeling process. Moreover, the inverted pendulum is fixed to the gears and motor, while the rocket is floating in the air, and thus canceling the gravity. The inverted pendulum has two more zeros in 0 and two real poles equidistant from zero, whereas the rocket has two poles in 0. The difference in poles is due to the absence of gravity into the rocket modeling, while the difference in zeros might be due to the fact that the arm is rigidly attached compared to the thruster.

Unfortunately even if the systems are close, an identical model could not be found. This difference might come from the setup and its extra arm, which introduce a rotational force instead of a horizontal force e.g an inverted pendulum with a cart.