\newcommand{\projectAbstract}{
This reports aims to investigate the inverted pendulum and rocket launches, as balancing an inverted pendulum and following a flight path with a rocket are both unstable systems that are controlled by moving the bottom part of a long and narrow cylinder. 

First a brief overview is given as to why the two systems have similar stability problems before a mathematical model is derived. Both models are nonlinear and are linearized and reduced to two simple models. It is found that the simple mathematical models for the two systems are not identical. A single controller controlling both systems therefore cannot be made. A controller for each system is thus designed and implemented on the respective setup. 

The controller designed for the inverted pendulum was implemented on an Arduino and uses two potentiometers and a tachometer as sensors. It is found that the controller balances the pendulum satisfactory to the specifications made. 

The controller designed for the rocket showed that it could follow the trajectory in a linear simulation based on the model. The rocket setup was built in its entirety by the group. The rocket controller was not successfully implemented due to time constraints.

In conclusion, while a rocket during flight and an inverted pendulum share similarities with instability, the models are not identical and a controller cannot be made to work with both systems. A controller showing satisfactory balancing of the inverted pendulum  was made, but the rocket controller ultimately was not tested.
}

\newcommand{\projectSynopsis}{
Synopsis}


