\input{chapters/Design/IPController.tex}
\chapter{Design of the Rocket and Gimbal Controller}
The following chapter describes the design of the rocket and its control system. The main objective is not to design the rocket, but to implement at control system that can stabilize it during launch and flight. 

\section{Rocket Design}
For the purpose of studying the problem of the rocket control, a model rocket was designed and built. Since mechanical design is out of this work's scope, the engineering of this rocket will not be explained, but the design CAD files will be available on the GitHub repository.\todo{Add the github link later.}
This design consists of 2 sections, that will be called "stages".


\textbf{First stage: propulsion}
\begin{itemize}[noitemsep]
	\item {A thruster / Solid Rocket Booster (SRB)}
	\item {A thrust vectoring mechanism, also known as gimbal, with two degrees of freedom}
	\item {Two servo motors for actuating the gimbal}
\end{itemize}

\textbf{Interstage:}
\begin{itemize}[noitemsep]
	\item {An empty fairing separating the propulsion stage from the electronics}
\end{itemize}
\textbf{Upper stage: control}
\begin{itemize}[noitemsep]
	\item A frame to contain the electronics.
	\item A PCB with a micro-controller.
	\item A plastic separator with anti-vibration bearings.
	\item On this separator is placed a gyroscope: the attitude sensor.
	\item A nose fairing.
\end{itemize}

\subsubsection*{Choice of thrusters for the rocket}
A thruster is a central component in all types of rocket. In the project the thruster will be chosen based on availability and lift force. The maximum weight of the rocket can not exceed 300 grams and the thruster should be able to lift this. The average thrust of the thruster should be have at least a average thrust of 3 Newton. The choice is limited to the thrusters which can be acquired within the European regulations. Through superficial research it is found that Klima 18 mm rocket motors is legal in all of Europe, and will be chosen for the thruster. The chosen thruster is the version D3-P with the specifications cf. table \ref{ThrusterValue}.

\begin{table}[]
\centering
\begin{tabular}{lll}
\hline
Parameter      & Value         & Unit \\ \hline
Total impulse  & 17,4          & [N]  \\
Average thrust & $\approx$ 3   & [N]  \\
Maximum thrust & $\approx$ 9 & [N]  \\
Burn duration  & $\approx$ 5,5 & [s]  \\
Weight         & 0,105         & [kg] \\
Length         & 0,07          & [m] 
\end{tabular}
\caption{DP-3 thruster specifications}
\label{ThrusterValue}
\end{table}
                
%http://www.modelrockets.co.uk/shop/klima-model-rocket-motors/d3-six-pack-18mm-rocket-glider-motor-p-3311.html

\subsubsection*{Physical Parameters of the Rocket}
The important factors for controlling the rocket is the physical parameters. These will effect how the rocket would transfer a input to its output. We can not control the force of the thruster, and this can will therefore not be a part of the important   		
\begin{table}[htbp]
	\centering
	\begin{tabular}{llll}
	\hline
	Piece & Parameter & Value & Unit \\ \hline
	Rocket$_{overall}$ & Height & 297 & {[}mm{]} \\
	Rocket$_{overall}$ & Weight & 0,28 & {[}kg{]} \\
	Interstage & Diameter & 67 & {[}mm{]} \\
	Thrust vectoring system & Max. angle & ? & {[}rad{]}\\
	Thrust vectoring system & Response time & ? & {[}rad/s{]}
	\end{tabular}
\caption{Parameters of the rocket.}
\label{Rocket_measurements}
\end{table}
\todo{Input the rest of the parameters.}

\startexplain
\explain{Rocket$_{overall}$ is the total weight of the system, including thruster, electronics, and rocket structure.}{}
\stopexplain


\subsection{Choice of Sensors for the Rocket}  \label{sec:Rocket_sensor_choice}

The following sections describes the sensors chosen for implementation in the rocket. It is chosen that the microcontroller unit, from here on named MCU, used in the system will be a Arduino Nano. The Nano is chosen based on its low weight (7 grams) and small size (18 x 45 mm) which will be an advantage when fitting it in the rocket. Further about the  is described cf. section \ref{sec:MCU}.    

As described in section \ref{sec:PRocketAnalysis} the rocket can be a system with instability problems. In the inverted pendulum these instabilities is detected trough sampling sensors and corrected trough a DC motor control system. The same parameters is considered when controlling the rocket. A form of sensor is needed to detect the orientation and position of the rocket, and a control system is needed to counteract changes from the initial trajectory.

Choosing sensors for the rocket will be weighted based on following parameters:

\begin{itemize}[noitemsep]
\item Compatibility.
\item Power consumption.
\item Availability. 
\item Physical dimensions and weight.
\end{itemize}

Needed is a sensor for measuring:
\begin{itemize}[noitemsep]
\item Orientation.
\item Acceleration.
\item Temperature and barometric pressure.
\end{itemize}

Determining the altitude, orientation and acceleration of the rocket can be done with different types of sensor. Two types of sensors can be considered when involving rockets; reference sensors and inertial sensors. Reference sensors have a external reference to measure from where inertial sensors measures changes in it physical state from its inertial state. Commonly used sensors and applications is listed, these determined from different types of rocket applications:

\begin{itemize}[noitemsep]
\item Gyroscope
\item Accelerometer
\item Global Positioning System (GPS)
\item Magnetometer
\item Barometric pressure sensor
\item Infrared camera
\item Solar panels
\end{itemize}

Some sensors can easily be declined from the choice. This is mainly because there application is most commonly with rocket going to higher altitudes or even in-orbit around earth, which is not the interest in this project. For example is the infrared camera implemented in some satellites and rocket to determine the position of the earth relative to the satellite. This will not be implemented in the rocket, because the altitude of the rocket is too low. Equally is solar panels implemented mostly in satellites, because the time span for change is lower than when launching and correcting a rocket.
Sensors considered in the project is gyroscopes, accelerometers, magnetometer/compass, GPS, barometric sensor. 

GPS is a relatively common used component in flying systems. It is implemented in many quad-copters, where the goal is to hold a position based on GPS. In the rocket it can be used to determined velocity, deviations from trajectory, and altitude. Precise and fast GPS units have a high cost and higher power consumption than alternatives. It is therefore decided that GPS will not be included as a component of choice.    

An accelerometer measures acceleration in one to three axis(x,y,z). The reference for measuring is the gravitational force. A single axis accelerometer can measure the acceleration in the direction it is orientated. And can for example be used to determine the velocity upwards flying rocket. This can as well be used to determine a travelled distance based on knowing the acceleration and time. In the case of flying a rocket a three-axis accelerometer would be the implemented, when considering that the rocket can move both lateral and vertical in its position.  


A gyroscope is on the other hand measuring the angular velocity changes in three dimensions. The difference between the accelerometer and gyroscope is that the gyroscope is capable of measuring the rate of rotation around an axis. It does not rely on a fixed reference and is commonly used in applications like drones and other flying objects. In the rocket it can be used to determine the orientation and rotation of the rocket based on measuring the rate of changes in any direction.  


Combining these gives a Inertial Measurement Unit(IMU), which is commonly used in model planes and quad-copters. The application of this is to obtain the objects position through measuring velocity, orientation, rotation with the gyroscope and accelerometer. Both types of components are dependent of temperature and barometric pressure, so often an IMU includes multiple types of additional sensors for calibration purposes. A choice is made to use an IMU, given that its application in similar types of flying systems verifies that it is suitable for a rocket.    	  
Some performance factors must be considered when choosing a IMU. For example is the g-force range of the IMU important. If the maximum ratings is lower than the acceleration of the rocket, then the sensor would not be able to give sufficient data at maximum acceleration. As well is the sensitivity of the accelerometer important. The rocket is a system with a high amplitude g-force when launching, and therefore a accelerometer with low sensitivity is preferable.

\subsubsection{Inertial Measurement Unit - GY-87}
GY-87\cite{web:GY80} is an IMU made available for use. It includes an MPU6050, which combines a 3-axis accelerometer and a 3-axis gyroscope, a BMP180 thermometer/barometer, and a HMC5883 3-axis magnetometer. It is chosen based on its combination of components, low power consumption of $\approx$ 6,5 mA in measurement mode. It is designed so it can be implemented with the Arduino Nano through I2C communication. 
I2C is a two wire protocol which is working trough serial communication. I2C is based on having two wires between the Arduino and the GY-87. The one wire SCL is the serial clock, and the other wire SDA is the data wire. All components on the GY-87 are convenient to implement with the Arduino. Further analysis is described cf. section \textbf{[ref]} \todo{Ref to to implementation part when it is done.}



\section{Rocket Controller Design}\label{sec:RocketControllerDesign}
The following sections explains the design procedure behind fitting the transfer function and dynamics of the rocket with a feedback controller. The controllers goal is to change the orientation of the thruster trough regulating the thrust vectoring system.    

\input{chapters/Design/rocket/rocket_design_controller.tex}

\subsection{Design of Thrust Vectoring System}
\todo{Raphael task describe gimbal. - Mathias}

  
